Avoiding accuracy-limiting pitfalls in the study of protein-ligand interactions with isothermal titration calorimetry

### 2015-07-10 manuscript/itc-worksheet.tex
Updated manuscript with information about where coefficients in error propagation come from.

Provided the relative errors in concentration $[X]_s$ are sufficiently small ($<$20\%) for the Taylor expansion above to be accurate, we can use Eq.~\ref{equation:relative-error} to estimate the relative error in the thermodynamic parameters $K$ and $\Delta H$ and site parameter $n$ given corresponding uncertainties from the least-squares fit ($s_\mathrm{K,LS}$, $s_\mathrm{\Delta H, LS}$, $s_\mathrm{n, LS}$) and the uncertainty in the titrant concentration $s_\mathrm{[X]_s}$~\cite{tellinghuisen-chodera:2011:anal-biochem:systematic-itc-errors},
\begin{eqnarray}
\left(\frac{s_{K}}{K}\right)^2 &=& \left(\frac{s_\mathrm{K,LS}}{K}\right)^2 + \left(\frac{s_\mathrm{[X]_s}}{[X]_s}\right)^2 \nonumber \\
\left(\frac{s_{\Delta H}}{\Delta H}\right)^2 &=& \left(\frac{s_\mathrm{\Delta H,LS}}{\Delta H}\right)^2 + \left(\frac{s_\mathrm{[X]_s}}{[X]_s}\right)^2  \nonumber \\
\left(\frac{s_{n}}{n}\right)^2 &=& \left(\frac{s_\mathrm{n,LS}}{n}\right)^2 + \left(\frac{s_\mathrm{[X]_s}}{[X]_s}\right)^2 . 
\label{equation:relative-error-in-thermodynamic-parameters}
\end{eqnarray}
Since the uncertainty in our [X]$_s$ is only 1\%, the 3\% LS fit uncertainty dominates for $K$; but for $\Delta H$ the titrant uncertainty is more important, increasing the RSE from 0.7\% to 1.2\%.  
These computations are automatically handled by the spreadsheet, which also computes $\Delta G$ and $\Delta S$ and their uncertainties.

@@

{\color{red}[JDC: To propagate the error in $n$, I also used the rule of adding in quadrature:
\begin{equation}
\left(\frac{s_n}{n}\right)^2 = \left(\frac{s_{n,\mathrm{LS}}}{n}\right)^2 + \left(\frac{s_\mathrm{[X]_s}}{[X]_s}\right)^2 .
\end{equation}
If this is inappropriate, I should change this---I haven't worked through this yet.]}

>>>>

Provided the relative errors in concentration $[X]_s$ are sufficiently small ($<$20\%) for the Taylor expansion above to be accurate, we can use Eq.~\ref{equation:relative-error} to estimate the relative error in the thermodynamic parameters $K$ and $\Delta H$ and site parameter $n$ given corresponding uncertainties from the least-squares fit ($s_\mathrm{K,LS}$, $s_\mathrm{\Delta H, LS}$, $s_\mathrm{n, LS}$) and in the titrant concentration $s_\mathrm{[X]_s}$~\cite{tellinghuisen-chodera:2011:anal-biochem:systematic-itc-errors},
\begin{eqnarray}
\left(\frac{s_{K}}{K}\right)^2 &=& \left(\frac{s_\mathrm{K,LS}}{K}\right)^2 + \left(\frac{s_\mathrm{[X]_s}}{[X]_s}\right)^2 \nonumber \\
\left(\frac{s_{\Delta H}}{\Delta H}\right)^2 &=& \left(\frac{s_\mathrm{\Delta H,LS}}{\Delta H}\right)^2 + \left(\frac{s_\mathrm{[X]_s}}{[X]_s}\right)^2  \nonumber \\
\left(\frac{s_{n}}{n}\right)^2 &=& \left(\frac{s_\mathrm{n,LS}}{n}\right)^2 + \left(\frac{s_\mathrm{[X]_s}}{[X]_s}\right)^2 . 
\label{equation:relative-error-in-thermodynamic-parameters}
\end{eqnarray}
where the terms $i$ and $j$ from Eq.~\ref{equation:relative-error} are determined from the manner in which [X]$_s$ influences the thermodynamic parameter of interest (see, e.g.~Table 1 of \cite{tellinghuisen:jpcb:2005:optimizing-experimental-parameters-itc}).
Since the uncertainty in our [X]$_s$ is only 1\%, the 3\% LS fit uncertainty dominates for $K$; but for $\Delta H$ the titrant uncertainty is more important, increasing the RSE from 0.7\% to 1.2\%.  
These computations are automatically handled by the spreadsheet, which also computes $\Delta G$ and $\Delta S$ and their uncertainties.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% SPREADSHEET FIGURE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
